---
title: "progress_report-favourforchu"
author: "Favour Forchu"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(rvest)
library(Lahman)
library(readr)

```

In this lab, I scraped Hall of Fame voting data for 2026 from Baseball Reference, cleaned the data, and compared it to the Lahman HOF table.


//Scrapping the data
```{r}
url <- "https://www.baseball-reference.com/awards/hof_2026.shtml"
html <- read_html(url)
tables <- html_table(html)
length(tables)
```

//Choose HOF table
```{r}
data <- tables[[1]]
head(data, 3)
```

//Fix Columns
```{r}
actual_col_names <- data[1, ]
colnames(data) <- actual_col_names
data <- data[-1, ]
head(data, 3)
```

//Date cleaning
```{r}
hof_2026 <- data %>% select(Name, Votes, `%vote`) %>% mutate(year = 2026, votes = as.numeric(Votes), percent_vote = readr::parse_number(`%vote`), inducted = ifelse(percent_vote >= 75, "Y", "N")) %>% select(candidate_name = Name, year, votes, percent_vote, inducted)

head(hof_2026)
```


//Save Clean data
```{r}
write_csv(hof_2026, "HallOfFame_2026.csv")
```

// Compare to lahman
```{r}
lahman_2026 <- HallOfFame %>% filter(yearID == 2026)
lahman_2026
```

```{r}
hof_2026 %>% count(inducted)
lahman_2026 %>% count(votedBy, inducted)
```

The scraped Baseball data and Lahman data don't match row-for-row because the two sources organize Hall of Fame information differently. Baseball Reference shows the full BBWAA ballot, while Lahman stores official voting records and may include different voting bodies.
